Modal Verification Patterns for Systems

### 2025-06-25 paper.tex
../..

\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
%\usepackage{t1enc}
%\usepackage[strings]{underscore}
%%%%%\usepackage{geometry}
\usepackage{wrapfig}
\usepackage[rawfloats=true]{floatrow} 
\restylefloat{figure}    
\DeclareMathSizes{4}{4}{4}{4}
\usepackage{varwidth}
\usepackage{lmodern}
\usepackage{amsfonts}
\usepackage[fleqn]{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{listings}
\usepackage{comment}
\usepackage{longtable}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[toc,page]{appendix}
\usepackage[export]{adjustbox}
\usepackage{amssymb}
\usepackage{graphics,graphicx}
\usepackage{mathpartir}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{arrows.meta,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc, shapes, backgrounds}
\usepackage{verbatim}
\usepackage{tikz-qtree}
\usepackage{pgfplots}
\usepackage{caption,capt-of}
\usepackage{color}
\usepackage{soul}
\setul{1ex}{0.8ex}
\definecolor{orange}{rgb}{1,0.5,0}
\setulcolor{orange}
\usepackage{subcaption}
\usepackage{lipsum}
\usepackage{afterpage}
\usepackage{fontawesome}
\usepackage[utf8]{inputenc}

@@

\subsection{Decomposing a System into its Constituents \emph{Contingently}}
\label{sec:definitions}
\begin{table*}[t]
\centering
\caption{Modal Decomposition of Program-Logics.}
\label{table:decomposition}
 %$\{\Diamond (P) \}$ write  $\{ P\}_{n}$
%$\{\overset{ICut^{n}}{\hookrightarrow} (P) \}$ return $\{ P\}_{n} $
\begin{tabular}{@{}lcccp{4cm}@{}}
  \toprule
{\scriptsize Modality}& {\scriptsize Context} & {\scriptsize Elements}  &  {\scriptsize Nominalization} & {\scriptsize Context Steps} \\ \midrule
\scriptsize Post-Crash~\cite{tejthesis,perennialgit,tejperennial19} & \scriptsize $\Diamond \; P $  & \scriptsize $  \ell \mapsto_{n}^{\overline{\gamma}} v $ & \scriptsize Strong &  \scriptsize Crash Recovery    \\ 
%GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
\scriptsize NextGen~\cite{larsnextgen25} &  \scriptsize $\overset{t}{\hookrightarrow} \; P$  &\scriptsize  Own (t(a))  & \scriptsize Strong & \scriptsize Determined Based on the Model$^{*}$   \\
\scriptsize StackRegion$^{*}$  ~\cite{larsnextgen25} & \scriptsize $\overset{ICut^{n}}{\hookrightarrow} \; P$      & \scriptsize $\fbox{n} \; \ell \mapsto v$                                 & \scriptsize Strong & \scriptsize Alloc and Return to/from stack\\
\scriptsize Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} & \scriptsize $\triangle_{\pi}$ and $\triangledown_{\pi}$      & \scriptsize $\ell \mapsto v$                                 & \scriptsize Weak & \scriptsize Fence Acquire and Release  \\
\scriptsize Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} & \scriptsize [r]P     &  \scriptsize $\ell \mapsto v$ & \scriptsize Weak                         & \scriptsize Address-Space Switch  \\ 
\scriptsize Ref-Count~\cite{amalreal2024}&\scriptsize @$_{\ell}$ P    & \scriptsize $\ell_1 \mapsto v$ & \scriptsize Weak  & \scriptsize Allocating, Dropping and Sharing a Reference \\ \bottomrule
\end{tabular}

{\scriptsize *The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25})}.
%and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
\end{table*}

@@

\section{Conclusion}
\label{sec:conclusion}
\todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
It also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). So not only are the modalities useful
for conducting proofs, the pieces we have identified seem to be an effective way of working out
a specific modality for a specific use-case.

As our main conclusion and future work, we think that these essential patterns in the designs of modal abstractions discussed constitute the fundamentals of \emph{verification patterns} when working out how to specify different kinds of systems code.
\bibliographystyle{splncs04}
\bibliography{software}
\newpage
\appendix

>>>>

\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
%\usepackage{t1enc}
%\usepackage[strings]{underscore}
%%%%%\usepackage{geometry}
\usepackage{wrapfig}
\usepackage[rawfloats=true]{floatrow} 
\restylefloat{figure}    
\DeclareMathSizes{4}{4}{4}{4}
\usepackage{varwidth}
\usepackage{lmodern}
\usepackage{amsfonts}
\usepackage[fleqn]{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{listings}
\usepackage{comment}
\usepackage{longtable}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[toc,page]{appendix}
\usepackage[export]{adjustbox}
\usepackage{amssymb}
\usepackage{graphics,graphicx}
\usepackage{mathpartir}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{arrows.meta,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc, shapes, backgrounds}
\usepackage{verbatim}
\usepackage{tikz-qtree}
\usepackage{pgfplots}
\usepackage{caption,capt-of}
\usepackage{color}
\usepackage{soul}
\setul{1ex}{0.8ex}
\definecolor{orange}{rgb}{1,0.5,0}
\setulcolor{orange}
\usepackage{subcaption}
\usepackage{lipsum}
\usepackage{afterpage}
\usepackage{fontawesome}
\usepackage[utf8]{inputenc}
\usepackage{rotating}

@@

\subsection{Decomposing a System into its Constituents \emph{Contingently}}
\label{sec:definitions}
\begin{sidewaystable*}[t]
\centering
\caption{Modal Decomposition of Program-Logics.}
\label{table:decomposition}
 %$\{\Diamond (P) \}$ write  $\{ P\}_{n}$
%$\{\overset{ICut^{n}}{\hookrightarrow} (P) \}$ return $\{ P\}_{n} $
\begin{tabular}{@{}lcccp{4cm}@{}}
  \toprule
{ Modality}& { Context} & { Elements}  &  { Nominalization} & { Context Steps} \\ \midrule
 Post-Crash~\cite{tejthesis,perennialgit,tejperennial19} &  $\Diamond \; P $  &  $  \ell \mapsto_{n}^{\overline{\gamma}} v $ &  Strong &   Crash Recovery    \\ 
%GC-Modality &      & \multicolumn{2}{c}{\multirow{2}{*}{Because why not}} & 8                                   & & ~\cite{}\\
 NextGen~\cite{larsnextgen25} &   $\overset{t}{\hookrightarrow} \; P$  &  Own (t(a))  &  Strong &  Determined Based on the Model$^{*}$   \\
 StackRegion$^{*}$  ~\cite{larsnextgen25} &  $\overset{ICut^{n}}{\hookrightarrow} \; P$      &  $\fbox{n} \; \ell \mapsto v$                                 &  Strong &  Alloc and Return to/from stack\\
 Memory-Fence~\cite{fsl,fsl++,derekrustbelt20} &  $\triangle_{\pi}$ and $\triangledown_{\pi}$      &  $\ell \mapsto v$                                 &  Weak &  Fence Acquire and Release  \\
 Address-Space~\cite{kuru2024modalabstractionsvirtualizingmemory} &  [r]P     &   $\ell \mapsto v$ &  Weak                         &  Address-Space Switch  \\ 
 Ref-Count~\cite{amalreal2024}& @$_{\ell}$ P    &  $\ell_1 \mapsto v$ &  Weak  &  Allocating, Dropping and Sharing a Reference \\ \bottomrule
\end{tabular}

{\scriptsize *The StackRegion Modality is an instance of NextGen (called the Independence Modality in \cite{larsnextgen25})}.
%and the $\Diamond$ Figure for Post-Crash Modality uses notationis taken from NextGen Modality paper ~\cite{larsnextgen25}.
\end{sidewaystable*}

@@

\section{Conclusion}
\label{sec:conclusion}
% \todo[inline]{this definitely needs adjusting for both audience and emphasis on this being a survey}
We have argued that there is a systematic pattern emerging in the use of modalities in the verification of low-level systems software,
and explained how a range of modern work fits into this pattern.
% We have argued the essential patterns that help to choose the modalities when working out how to specify different kinds of systems code based on recent successes.
While we have focused our attention on public verification work carried out in Iris,
it also captures (a slightly more organized version of) our own thinking in coming up with designs for specifying distributed systems~\cite{gordon2019modal}, 
% virtual memory managers~\cite{kuru2024modalabstractionsvirtualizingmemory}, 
and in ongoing work using different modalities for different parts of a copy-on-write filesystem (e.g., for assertions true in different snapshots). 
So not only are the modalities observably present, but they are demonstrably useful
for conducting proofs of systems software.
%  the pieces we have identified seem to be an effective way of working out
% a specific modality for a specific use-case.
We believe that these essential patterns in the designs of modal abstractions for systems verification constitute a fundamental
concept
when working out how to specify different kinds of systems code.

\bibliographystyle{splncs04}
\bibliography{software}
\newpage
\appendix
